Region-network hierarchical sparsity priors for\\ high-dimensional inference in brain imaging

### 2016-01-30 paper/paper.tex
big pass over sections

\begin{abstract}
Structured sparsity penalization has recently improved
statistical models applied to high-dimensional data in various domains.
As an extension to imaging neuroscience,
priors on network hierarchies of brain regions have been incorporated into
logistic-regression estimators for neural activity maps.
This remarries the perspectives of
functional segregation and functional integration
that are routinely divorced in neuroscientific research.
%
Region-network hierarchical priors for supervised classification
of 18 psychological tasks
from a reference dataset are
shown to outperform na\"ive $\ell_1$-norm and
trace-norm penalization, as well as neurobiologically
informed $\ell_1/\ell_2$-block-norm group sparsity.
Varying the relative importance of
region and network structure within the hierarchical tree prior
recovered complementary aspects of the neural activity pattern
with similar predictive accuracies.
%
In sum,
we demonstrate that priors of local and global neurobiological knowledge can enhance
out-of-sample performance and domain interpretability
by hierarchical tree sparsity.

@@

% specialization & integration
Concepts on human brain organization have long been torn
between the two extremes
\textit{functional specialization} and \textit{functional integration}.
Functional specialization emphasizes that microscopically distinguishable
brain regions are responsible distinct classes of computational processes
\cite{kanwisher2010functional}.
Functional integration, in turn, emphasizes that brain function
is enabled by complex connections between these
distinct brain regions \cite{sporns14nn}.
%
These notions were predominantly derived from
invasive examination of anatomy (i.e., histological preparation),
connectivity (i.e., axonal tracing),
and functional properties
(i.e., single-cell recordings) in animals.
Regarding functional segregation into specialized regions,
early histological investigations into the microscopic heterogeneity of
the human cerebral cortex have resulted
in several detailed anatomical maps
\cite{brodmann1909vergleichende, vogt1919allgemeine}.
Regarding axonal connections,
each such cortical area has been observed
to possess a unique set of incoming and outgoing connections
\cite{passingham2002, young93monkey, scannell95cat}.
%
Both
local
%cyto- and chemoarchitectonic
infrastructure
and its unique global connectivity profile
together are thought to realize brain function.
%
In sum,
cortical brain modules versus connections between them
reflect 
functional specialization versus functional integration
\cite{friston2002beyond, mesulam_sensation}.
Importantly,
probably no existing brain analysis method acknowledges that
both functional organizations are inextricably involved
in the realization of mental operations
\cite{tononi1998complexity, saygin2012}.

@@

% study
Building on these two major interpretational streams in systems neuroscience,
the present study proposes to incorporate
established neurobiological structure underlying
functional segregation and integration
into supervised estimators
by hierarchical structured sparsity.
% The "true" relative importance of 
% local region compartments and global network compartments
% is typically unknown
% but
% probably varying in degree
% across diverse neuroscientific questions.
%
Learning techniques
exploiting structured sparsity 
have recently made much progress in various application domains
from processing of auditory signals \cite{daudet2004sparse},
natural images \cite{harzallah2009combining} and
videos \cite{kang2015structured, kim2010sparse}
to
astrophysics \cite{vinci2014estimating},
genetics \cite{rapaport2008classification, kim2012tree},
and
conformational dynamics of protein complexes \cite{jenatton2009structured}.
%
This is extended by the present work that
introduced neuroscience-specific estimators capitalizing on
neurobiologically plausible region and network priors.
%
Based on the largest neuroimaging repository,
we demonstrated that domain-informed supervised models
gracefully tackle the curse of dimensionality,
yield more human-interpretable results,
and generalize better to new samples
than domain-na\"ive estimators.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/sparsities.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Model performance across sparsity priors.}
Compares the performance of logistic regression estimators
with 4 different structured and unstructered sparsity terms
in classifying neural activity from 18 psychological tasks.
The class-wise precision and recall metrics
were obtained on the same test set.
%
Unstructured $\ell_1$-penalized logistic regression (\textit{upper left})
imposed a minimum of relevant brain voxels without
assuming special structure.
Structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
(\textit{lower left})
imposed region compartments, but na\"ive to network structure.
Structured trace-norm penalization (\textit{lower right})
imposed low-rank structure
with sparsity of network patterns, but na\"ive to region structure.
Structured $\ell_1/\ell_2$ block norm with a hierarchy of
both region and network priors (\textit{upper right})
exhibited the best out-of-sample performance.
%
A priori knowledge of both region and network neighborhoods
was hence most beneficial for predicting psychological tasks from
brain maps.
}
\label{fig_sparsities}
\end{figure}

\paragraph{Benchmarking hierarchical tree sparsity against
common sparsity penalties.}
Hierarchical region-network priors have been systematically
evaluated against other popular choices of sparse classification algorithms
in an 18-class scenario
(Figure \ref{fig_sparsities}).
%
Logistic regression with $\ell_1/\ell_2$ block norm penalization
incorporated a hierarchy of previously known region and network neighborhoods
for a neurobiological bias of the statistical estimation.
%
Vanilla logistic regression with $\ell_1$-penalization
does not assume any previously known special structure.
This classification estimator embraces a vision of neural activity structure
that expects a minimum of
topographically and functionally independent brain voxel to be relevant.
%
Logistic regression with sparse group sparsity
imposes a structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
with a known atlas of region voxel groups onto the statistical estimation process.
This supervised estimator shrinks and selects the coefficients
of topographically compact voxel groups expected to be relevant together.
%
Logistic regression with trace-norm penalization
imposed low-rank structure.
This supervised classification algorithm
expected a minimum of unknown network patterns
to be relevant.
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Hierarchial tree sparsity demonstrated the best generalization
in distinguishing unseen neural activity maps from 18 tasks
(89.67\%, mean recall XX.YY, mean precision XX.YY).
It was closely followed by logistic regression with trace-norm
regularization that is structured but not neurobiologically informed
(89.43\%, mean recall XX.YY, mean precision XX.YY).
Lasso featured an average performance comparing to the other sparse estimators
(88.60\%, mean recall XX.YY, mean precision XX.YY)
Introduced a priori knowledge of brain region compartments
by sparse group sparsity performed worst
(87.89\%, mean recall XX.YY, mean precision XX.YY).
%
In an important subanalysis,
the gain of the combined region-network prior was confirmed by
selectively zeroing the $eta_g$ coefficients of either all region groups
or all network groups in the hierarchical prior.
Removing region structure from the prior achieved
88,84\% accuracy (mean recall XX.YY, mean precision XX.YY),
while removing network structure from the prior achieved
87,05\% accuracy (mean recall XX.YY, mean precision XX.YY).
These results from partial priors are indeed outperformed
the full region-network tree prior at 89,67\% accuracy
(mean recall XX.YY, mean precision XX.YY).
%
In sum,
biasing sparse model selection by domain knowledge of region-network hierarchies
outperformed other types of frequently used sparse penalization techniques.

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/dataratio.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Na\"ive versus informed sparse model selection
across training set sizes.}
Ordinary $\ell_1$-penalized logistic regression
(\textit{upper row})
is compared
to hierarchical-tree-penalized logistic regression
(\textit{lower row})
with increasing fraction
of the available training data (\textit{left to right columns}).
For one example (i.e., ``View tools'') from 18 psychological tasks,
unthresholded axial maps of model weights
are shown for comparison against
the sample average of that class
(\textit{rightmost column}, thresholded at the $75^{th}$ percentile).
The out-of-sample accuracies for predicting all 18 classes
is given in percent.
%
In the data-scarce scenario,
typical for brain imaging,
hierarchical tree sparsity achieves much
better support recovery with the biggest difference
in model performance.
%
In the data-rich scenario,
neurobiologically informed logistic regression
profits more from the available information quantities than
neurobiologically naive logistic regression.
}
\label{fig_dataratio}
\end{figure}

\paragraph{Sample complexity of na\"ive versus informed sparse model selection.}
Subsequently, the sample complexity of
$\ell_1$-penalized and hierarchical-tree-penalized logistic regression
were quantitatively compared (Figure \ref{fig_dataratio}).
Region-network priors should bias model selection towards more
neurobiologically plausible classification estimators.
This should yield better out-of-sample generalization and
support recovery than
$\ell_1$-constrained logistic regression na\"ive to neurobiology
in the data-scarce and data-rich scenarios.
%
The HCP task data with examples from 18 classes were first divided into
90\% of training set (i.e., 7584 neural activity maps) and
10\% of test set (i.e., 842 maps).
Both learning algorithms were fitted based on the
training set at different subsampling fractions:
20\% (1516 maps),
40\% (3033 maps),
60\% (4550 maps),
80\% (6067 maps), and
100\% (7584 maps).
%
The stratified and shuffled training data were submitted to
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
%
Three observation have been made.
In the data-scarce scenario (i.e., 1/5 of actual training data),
hierarchical tree sparsity achieved the biggest advantage
in out-of-sample performance by 5,11\% as well as
better support recovery with weight maps already much closer
to the training data average.
In the case of scarce training data, which is typical for the brain imaging domain,
regularization by region-network priors indeed allowed for
more effective extraction of classification-relevant structure
from the neural activity scans.
%
Across scenarios,
the weight maps from ordinary logistic regression exhibit
higher variance and many more zero coefficients
than hierarchical tree logistic regression.
Given the usually high multicollinearity in neuroimaging data,
this observation is likely to reflect instable selection of
representatives among class-responsive predictor groups
due to the $\ell_1$-norm penalization.
%
In the data-rich scenario (i.e., entire training data used for model fitting),
neurobiologically informed logistic regression
profits more from the increased information quantities than
neurobiologically naive logistic regression.
That is, the region-network priors actually further enhance the similarity
to the weight maps even in abundant input data.
This was the case although
the maximal classification performance of $\approx$90\% has already
been reached with small training data fractions by the structured estimator.
In contrast, 
the unstructured estimator reached this generalization performance
only with bigger input data quantities.

@@

\paragraph{Support recovery as a function of region and network emphasis.}
Finally, the relative importance of the
region and network priors within the hierarchical tree prior
was quantified (Figure \ref{fig_regnetratio}).
The $\eta_g$ group of region priors was multiplied with a
region-network ratio, while the
$\eta_g$ group of network priors was biased by the corresponding
network-region ratio. A region-network ratio of 3, for instance,
increased the relative importance of known region structure
by multiplying $\frac{3}{1}$ to the
$\eta_g$ factor of all region groups
and multiplying
$\frac{1}{3}$ to all network groups.
The data splitting cross-validation scheme was identical to the
above modelling experiments.
%
As the most important observation,
a range between region-dominant and network-dominant structured penalties
yields quantitatively almost identical generalization to new data
but qualitatively different decision functions manifested in the weight maps
(Figure \ref{fig_regnetratio}, second and forth column).
Classification models with many zero coefficients but high absolute
coefficients in either region compartments or network compartments
can similarly extrapolation to unseen neural activity maps.
Second,
these perform similar to equilibrated region-network priors
that set less voxel coefficients to zero and spread the
probability mass with lower absolute coefficients across the whole brain
(Figure \ref{fig_regnetratio}, third column in the middle).
Third,
overly strong emphasis on either level of the hierarchical prior
can yield the neurobiologically informative maps
of the most necessary region or network structure for
statistically significant out-of-sample performance
(Figure \ref{fig_regnetratio}, leftmost and rightmost columns).
%
In sum,
stratifying the hierarchical tree penalty between region and network emphasis
suggests that \textit{class-specific region-network weights}
might offer more performant and more interpretable classification models
in the future.

\section{Discussion}
Relevant structure in neuroimaging data has long been investigated
according to two separate organizational principles:
functional segregation into discrete brain regions
and functional integration by interregional brain networks.
%
This paper demonstrates the simultaneous exploitation of
both these neurobiological compartments
for sparse variable selection and high-dimensional prediction
in a reference dataset.
%
Introducing existing domain knowledge into model selection
allowed privileging members of the function space
that are most neurobiologically plausible.
%
Domain-informed hierarchical structured sparsity is shown to enhance
both model interpretability and generalization performance,
although these statistical-learning goals are typically in conflict.

The present approach has important advantages over previous
neuroimaging studies that capitalized on dimensionality reduction to harness
the curse of dimensionality.
They often used preliminary region-wise pooling functions
or regression against network templates
for subsequent supervised learning on the aggregated feature space.
Such lossy two-step approaches of feature engineering and inference
\textit{i)} can only account for either functional specialization or
functional integration of brain organization,
\textit{ii)} depend on the ground truth being a region or network effect,
and
\textit{iii)} cannot issue individual coefficients for every brain voxels.
%
Hierarchical region-network sparsity overcomes these shortcomings
by estimating individual voxel contributions
while benefitting from their functional segregation and integration
to restrict statistical complexity.
%
Viewed from the bias-variance tradeoff,
our modification to logistic regression estimators
entailed a large decrease in model variance but only a modest
increase in model bias.
Viewed from the Vapnik-Chervonenkis dimensions,
this entailed a healthy decrease in the complexity capacity of the prediction model
with a higher chance of generalizing to unobserved data.

In the future,
region-network sparsity priors could be incorporated into various
pattern-learning methods in systems neuroscience.
%
This includes supervised methods for whole-brain classification and regression
with one or several target variables.
The principled regularization scheme could even inform
unsupervised structure discovery methods,
such as principal component analysis
\cite{jenatton2009structured}
and
k-means clustering \cite{witten2010framework}.
%
Additionally,
model regularization by hierarchical structured sparsity could be extended
from the spatial domain of neural activity to
priors of coherent spatiotemporal activity structure
\cite{gramfort2011tracking}.
%
Ultimately,
successful high-dimensional inference is
an important prerequisite
for predicting diagnosis,
disease trajectories, and treatment responses
in personalized psychiatry and neurology.

>>>>

\begin{abstract}
Structured sparsity penalization has recently improved
statistical models applied to high-dimensional data in various domains.
As an extension to imaging neuroscience,
priors on network hierarchies of brain regions have been incorporated into
logistic-regression estimators for neural activity maps.
This remarries the perspectives of
functional segregation and functional integration
that are routinely divorced in neuroscientific research.
%
Region-network hierarchical priors for supervised classification
of 18 psychological tasks
from a reference dataset are
shown to outperform na\"ive $\ell_1$-norm and
trace-norm penalization, as well as neurobiologically
informed $\ell_1/\ell_2$-block-norm group sparsity.
Varying the relative importance of
region and network structure within the hierarchical tree prior
recovered complementary aspects of the neural activity patterns
with similar predictive accuracies.
%
In sum,
we demonstrate that priors of local and global neurobiological knowledge can enhance
out-of-sample performance and domain interpretability
by hierarchical tree sparsity.

@@

% specialization & integration
Concepts on human brain organization have long been torn
between the two extremes
\textit{functional specialization} and \textit{functional integration}.
Functional specialization emphasizes that microscopically distinguishable
brain regions are responsible distinct classes of computational processes
\cite{kanwisher2010functional}.
Functional integration, in turn, emphasizes that brain function
is enabled by complex connections between these
distinct brain regions \cite{sporns14nn}.
%
These notions were predominantly derived from
invasive examination of anatomy (i.e., histological preparation),
connectivity (i.e., axonal tracing),
and functional properties
(i.e., single-cell recordings) in animals.
Regarding functional segregation into specialized regions,
early histological investigations into the microscopic heterogeneity of
the human cerebral cortex have resulted
in several detailed anatomical maps
\cite{brodmann1909vergleichende}.
Regarding axonal connections,
each such cortical area has been observed
to possess a unique set of incoming and outgoing connections
\cite{passingham2002, young93monkey, scannell95cat}.
%
Both
local
%cyto- and chemoarchitectonic
infrastructure
and its unique global connectivity profile
together are thought to realize brain function.
%
In sum,
cortical brain modules versus connections between them
reflect 
functional specialization versus functional integration
\cite{friston2002beyond, mesulam_sensation}.
Importantly,
probably no existing brain analysis method acknowledges that
both functional organizations are inextricably involved
in the realization of mental operations
\cite{tononi1998complexity, saygin2012}.

@@

% study
Building on these two major interpretational traditions in systems neuroscience,
the present study proposes to incorporate
established neurobiological structure underlying
functional segregation and integration
into supervised estimators
by hierarchical structured sparsity.
% The "true" relative importance of 
% local region compartments and global network compartments
% is typically unknown
% but
% probably varying in degree
% across diverse neuroscientific questions.
%
Learning techniques
exploiting structured sparsity 
have recently made much progress in various application domains
from processing auditory signals \cite{daudet2004sparse},
natural images \cite{harzallah2009combining} and
videos \cite{kang2015structured, kim2010sparse}
to
astrophysics \cite{vinci2014estimating},
genetics \cite{rapaport2008classification, kim2012tree},
and
conformational dynamics of protein complexes \cite{jenatton2009structured}.
%
This is extended by the present work that
introduced neuroscience-specific estimators capitalizing on
neurobiologically plausible region and network priors.
%
Based on the largest neuroimaging repository,
we demonstrated that domain-informed supervised models
gracefully tackle the curse of dimensionality,
yield more human-interpretable results,
and generalize better to new samples
than domain-na\"ive estimators.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/sparsities.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Model performance across sparsity priors.}
Compares the performance of logistic regression estimators
with 4 different structured and unstructered sparsity terms
in classifying neural activity from 18 psychological tasks.
The class-wise precision and recall metrics
were obtained on the same test set.
%
Unstructured $\ell_1$-penalized logistic regression (\textit{upper left})
imposed a minimum of relevant brain voxels without
assuming special structure.
Structured $\ell_1/\ell_2$-block norm with additional $\ell_1$ term
(\textit{lower left})
imposed region compartments, but na\"ive to network structure.
Structured trace-norm penalization (\textit{lower right})
imposed low-rank structure
with sparsity of network patterns, but na\"ive to region structure.
Structured $\ell_1/\ell_2$-block-norm with a hierarchy of
both region and network priors (\textit{upper right})
exhibited the best out-of-sample performance.
%
A priori knowledge of both region and network neighborhoods
was hence most beneficial for predicting psychological tasks from
brain maps.
}
\label{fig_sparsities}
\end{figure}

\paragraph{Benchmarking hierarchical tree sparsity against
common sparsity penalties.}
Hierarchical region-network priors have been systematically
evaluated against other popular choices of sparse classification algorithms
in an 18-class scenario
(Figure \ref{fig_sparsities}).
%
Logistic regression with $\ell_1/\ell_2$ block norm penalization
incorporated a hierarchy of previously known region and network neighborhoods
for a neurobiological bias of the statistical estimation.
%
Vanilla logistic regression with $\ell_1$-penalization
does not assume any previously known special structure.
This classification estimator embraces a vision of neural activity structure
that expects a minimum of
topographically and functionally independent brain voxel to be relevant.
%
Logistic regression with sparse group sparsity
imposes a structured $\ell_1/\ell_2$ block norm with additional $\ell_1$ term
with a known atlas of region voxel groups onto the statistical estimation process.
This supervised estimator shrinks and selects the coefficients
of topographically compact voxel groups expected to be relevant together.
%
Logistic regression with trace-norm penalization
imposed low-rank structure.
This supervised classification algorithm
expected a minimum of unknown network patterns
to be relevant.
%
The stratified and shuffled training data were submitted
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.

Across analyses,
hierarchial tree sparsity was most successful
in distinguishing unseen neural activity maps from 18 psychological tasks
(89.67\%, mean recall XX.YY, mean precision XX.YY).
It was closely followed by logistic regression
structured by trace-norm regularization
(89.43\%, mean recall XX.YY, mean precision XX.YY).
Lasso featured an average performance comparing to the other sparse estimators
(88.60\%, mean recall XX.YY, mean precision XX.YY).
Introducing a priori knowledge of brain region compartments
by sparse group sparsity performed worst
(87.89\%, mean recall XX.YY, mean precision XX.YY).
%
In an important subanalysis,
the gain of the combined region-network prior was also confirmed by
selectively zeroing the $\eta_g$ coefficients of all region groups
or all network groups in the hierarchical prior.
Removing region structure from the prior achieved
88,84\% accuracy (mean recall XX.YY, mean precision XX.YY),
while removing network structure from the prior achieved
87,05\% accuracy (mean recall XX.YY, mean precision XX.YY).
These results from partial priors are indeed outperformed
the full region-network tree prior at 89,67\% accuracy
(mean recall XX.YY, mean precision XX.YY).
%
In sum,
biasing sparse model selection by domain knowledge of region-network hierarchies
outcompeted other types of frequently used sparse penalization techniques.

\begin{figure}
\begin{centering}
\includegraphics[width=1.00\textwidth]{figures/dataratio.pdf}
\end{centering}
\vspace{-0.6cm}
\caption{\textbf{Na\"ive versus informed sparse model selection
across training set sizes.}
Ordinary $\ell_1$-penalized logistic regression
(\textit{upper row})
is compared
to hierarchical-tree-penalized logistic regression
(\textit{lower row})
with increasing fraction
of the available training data (\textit{left to right columns}).
For one example (i.e., ``View tools'') from 18 psychological tasks,
unthresholded axial maps of model weights
are shown for comparison against
the sample average of that class
(\textit{rightmost column}, thresholded at the $75^{th}$ percentile).
The out-of-sample accuracies for predicting all 18 psychological tasks
is given in percent.
%
In the data-scarce scenario,
typical for brain imaging,
hierarchical tree sparsity achieves much
better support recovery with the biggest difference
in model performance.
%
In the data-rich scenario,
neurobiologically informed logistic regression
profits more from the available information quantities than
neurobiologically naive logistic regression.
}
\label{fig_dataratio}
\end{figure}

\paragraph{Sample complexity of na\"ive versus informed sparse model selection.}
Subsequently, the sample complexity of
$\ell_1$-penalized and hierarchical-tree-penalized logistic regression
were quantitatively compared (Figure \ref{fig_dataratio}).
Region-network priors should bias model selection towards more
neurobiologically plausible classification estimators.
This should yield better out-of-sample generalization and
support recovery than
neurobiology-na\"ive $\ell_1$-constrained logistic regression
in the data-scarce and data-rich scenarios.
%
The HCP task data with examples from 18
psychological tasks were first divided into
90\% of training set (i.e., 7584 neural activity maps) and
10\% of test set (i.e., 842 neural activity maps).
Both learning algorithms were fitted based on the
training set at different subsampling fractions:
20\% (1516 maps),
40\% (3033 maps),
60\% (4550 maps),
80\% (6067 maps), and
100\% (7584 maps).
%
The stratified and shuffled training data were submitted
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.

Three observations have been made.
In the data-scarce scenario (i.e., 1/5 of available training data),
hierarchical tree sparsity achieved the biggest advantage
in out-of-sample performance by 5,11\% as well as
better support recovery with weight maps already much closer
to the class averages.
In the case of scarce training data, which is typical for the brain imaging domain,
regularization by region-network priors thus allowed for
more effective extraction of classification-relevant structure
from the neural activity maps.
%
Across training data fractions,
the weight maps from ordinary logistic regression exhibited
higher variance and more zero coefficients
than hierarchical tree logistic regression.
Given the usually high multicollinearity in neuroimaging data,
this observation is likely to reflect instable selection of
representatives among class-responsive predictor groups
due to the $\ell_1$-norm penalization.
%
In the data-rich scenario (i.e., entire training data used for model fitting),
neurobiologically informed logistic regression
profited more from the increased information quantities than
neurobiologically na\"ive logistic regression.
That is, the region-network priors actually further enhance the similarity
to the weight maps even in abundant input data.
This was the case although
the maximal classification performance of $\approx$90\% has already
been reached with small training data fractions by the structured estimator.
In contrast, 
the unstructured estimator reached this generalization performance
only with bigger input data quantities.

@@

\paragraph{Support recovery as a function of region and network emphasis.}
Finally, the relative importance of the
region and network priors within the hierarchical tree prior
was quantified (Figure \ref{fig_regnetratio}).
The group weight $\eta_g$ of region priors was multiplied with a
region-network ratio, while the
group weight $\eta_g$ of network priors was divided by that
region-network ratio. For instance, a region-network ratio of 3
increased the relative importance of known region structure
by multiplying $\frac{3}{1}$ to
$\eta_g$ of all region group penalties
and multiplying
$\frac{1}{3}$ to $\eta_g$ of all network group penalties.
The data splitting cross-validation scheme was identical to the
above modelling experiments.

As the most important observation,
a range between region-dominant and network-dominant structured penalties
yielded quantitatively almost identical generalization to new data
but qualitatively different decision functions manifested in the weight maps
(Figure \ref{fig_regnetratio}, second and forth column).
Classification models with many zero coefficients but high absolute
coefficients in either region compartments or network compartments
can similarly extrapolate to unseen neural activity maps.
Second,
these achieve classification performance
comparable to equilibrated region-network priors
that set less voxel coefficients to zero and spread the
probability mass with lower absolute coefficients across the whole brain
(Figure \ref{fig_regnetratio}, third column in the middle).
Third,
overly strong emphasis on either level of the hierarchical prior
can yield the neurobiologically informative results with maps
of the most necessary region or network structure for
statistically significant out-of-sample performance
(Figure \ref{fig_regnetratio}, leftmost and rightmost columns).
%
In sum,
stratifying the hierarchical tree penalty between region and network emphasis
suggests that \textit{class-specific region-network weights}
might offer more performant and more interpretable classification models
in the future.

\section{Discussion}
Relevant structure in neuroimaging data has long been investigated
according to two separate organizational principles:
functional segregation into discrete brain regions
\cite{passingham2002}
and functional integration by interregional brain networks
\cite{sporns14nn}.
%
This proof-of-concept study demonstrates the simultaneous exploitation of
both these neurobiological compartments
for sparse variable selection and high-dimensional prediction
in a reference dataset.
%
Introducing existing domain knowledge into model selection
allowed privileging members of the function space
that are most neurobiologically plausible.
%
Domain-informed hierarchical structured sparsity is shown to enhance
both model interpretability and generalization performance,
although these statistical-learning goals are typically in conflict.

The present approach has important advantages over previous
analysis strategies that rely on dimensionality reduction of
the neuroimaging data to harness the curse of dimensionality.
They often use preliminary pooling functions within regions
or regression against network templates
for subsequent supervised learning on the aggregated feature space.
Such lossy approaches divided into feature engineering and inference steps
\textit{i)} can only satisfy the specialization or
integration account of brain organization,
\textit{ii)} depend on the ground truth being a region or network effect,
and
\textit{iii)} cannot issue individual coefficients for every brain voxels.
%
Hierarchical region-network sparsity addresses these shortcomings
by estimating individual voxel contributions
while benefitting from their functional segregation and integration
to restrict statistical complexity.
%
Viewed from the bias-variance tradeoff,
our modification to logistic regression estimators
entailed a large decrease in model variance but only a modest
increase in model bias.
Viewed from the Vapnik-Chervonenkis dimensions,
this entailed a healthy decrease in the complexity capacity of the prediction model
with a higher chance of generalizing to unobserved data.

In the future,
region-network sparsity priors could be incorporated into various
pattern-learning methods in systems neuroscience.
%
This includes supervised methods for whole-brain classification and regression
with one or several target variables.
The principled regularization scheme could even inform
unsupervised structure-discovery methods,
such as principal component analysis
\cite{jenatton2009structured}
and
k-means clustering \cite{witten2010framework}.
%
Additionally,
model regularization by hierarchical structured sparsity could be extended
from the spatial domain of neural activity to
priors of coherent spatiotemporal activity structure
\cite{gramfort2011tracking}.
%
Ultimately,
successful high-dimensional inference is
an important prerequisite
for predicting diagnosis,
disease trajectories, and treatment response
in personalized psychiatry and neurology.
